\documentclass{article}
\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{comment}
\usepackage{eurosym}

\graphicspath{ {plots/} }

\title{Dynamic Optimization - Homework assignment 7}
\author{Christian Segercrantz 481056}


\begin{document}
	\maketitle
	\pagebreak

\section*{7.6}

We know that on the 0th and 1st day the probabilities of the demand being 0, 1, 2, or 3 is $20\%, 40\%, 30\% \text{ and } 10\%$ and after that $10\%, 20\%, 30\%, \text{ and } 40\%$ respectively. We can sell a package at 5\euro รก piece. We additionally know that we can manufacture 0-3 packages per day at the cost of 0\euro, 4\euro, 7\euro, or 10\euro. We also assume that unsold packages can transfer to the following day. The mean of the demand of the 0th and 1st day is 1.3 and for the remaining days 2. We assume that $x_0 = 0$ and that since the demand is 0 in the end $J_4(x_4) = 0$

We can model the packages per day as
\begin{equation}
	x_{k+1} = x_{k} + u_k - w_k, \qquad x_{k+1}\geq 0
\end{equation}
where x is the amount of items per day, u is the amount of new packages fabricated and w is the stochastic demand. Here we assume that the packs can be stored from one day to the next. 

The cost can be modeled as 
\begin{equation}
	g_k(x_k,u_k,w_k) = c u_k - 5\min(w_k,x_k) =
	 \begin{cases}
		c u_k - 5x_k, & w_k \geq x_k\\
		c u_k - 5w_k, & w_k < x_k
	\end{cases} 
\end{equation}
where c describes the cost per fabricated package.

The control policy $u_k = \mu^*_k(x_k)$ has the following constrains
\begin{equation}
	u \in
	\begin{cases}
		\lbrace0,1,2,3\rbrace, &x_k = 0\\
		\lbrace0,1,2\rbrace, &x_k = 1\\
		\lbrace0,1\rbrace, &x_k = 2\\
		\lbrace0\rbrace, &x_k = 3\\	
	\end{cases}
\end{equation}

We want to minimize the cost  according to the formula
\begin{equation}
	\sum_{n=1}^{3} g_k(x_k,u_k,w_k)
\end{equation}
where c is the cost of production.

\begin{align}
	J_k(x_k) &= \max_{u_k} \mathbb{E}_{w_k} \lbrace g_k(x_k,u_k,w_k) + J_{k+1}(x_{k+1}) \rbrace \\
	&= \max_{u_k} \mathbb{E}_{w_k} \lbrace g_k(x_k,u_k,w_k) + J_{k+1}(x_{k} + u_k - w_k) \rbrace
\end{align}
\subsection*{a)}
We start with k=3
\begin{align}
	J_3(x_3) =& 
	\begin{cases}
		c u_3 - 5x_3 + J_4(x_4), & w_3 \geq x_3\\
		c u_3 - 5w_3 + J_4(x_4), & w_3 < x_3
	\end{cases}\\
	\mathbb{E}_{w_k}\lbrace J_3(x_3)\rbrace =& cu_3 -5(0.1\min(0,\max(x_3+u_3,3)) + 0.2\min(1,\max(x_3+u_3,3)) +\\ &0.3\min(2,\max(x_3+u_3,3))+0.4\min(3,\max(x_3+u_3,3)))
\end{align}
We can write a table to find the optimal control policy which can be seen in Table \ref{tab:k3}. An example of the calculation is e.g. for $x_3 = 2, u_3 = 1$: $4-5(0.1\cdot0 + 0.2\cdot1 +0.3\cdot2+0.4\cdot3) = -6$
\begin{table}[h]
	\centering
	\caption{$k=3$, the minimi per row is marked with bold.}
	\label{tab:k3}
	\begin{tabular}{l|llll|l}
		k=3     & $u_3=0$       & $u_3=1$ & $u_3=2$     & $u_3=3$ & $\mu_3^*$ \\ \hline
		$x_3=0$ & 0             & -0.5    & \textbf{-1} & 0       & 2         \\
		$x_3=1$ & \textbf{-4.5} & -4      & -3          & -       & 0         \\
		$x_3=2$ & \textbf{-8}   & -6      & -           & -       & 0         \\
		$x_3=3$ & \textbf{-10}  & -       & -           & -       & 0        
	\end{tabular}
\end{table}

For k=2 we then get
\begin{align}
	J_2(x_2) = &
	\begin{cases}
		c u_2 - 5x_2 + J_3(x_3), & w_2 \geq x_2\\
		c u_2 - 5w_2 + J_3(x_3), & w_2 < x_2
	\end{cases}\\
	=& \begin{cases}
		c u_2 - 5x_2 + J_2(x_{2} + u_2 - w_2), & w_2 \geq x_2\\
		c u_2 - 5w_2 + J_2(x_{2} + u_2 - w_2), & w_2 < x_2
	\end{cases}\\
	J_2(x_2) =& \mathbb{E}_{w_k}\lbrace g_2(x_2,u_2,w_2)\rbrace + \mathbb{E}_{w_k}\lbrace J_3(x_{2} + u_2 - w_2)\rbrace.
\end{align}
The first term is calculated the same way as earlier, but the second term is now the expected value of the optimal values from earlier. I.e. for $x_2=2, u_2=1$ we get that 
\begin{equation}
	\mathbb{E}_{w_k}\lbrace J_3(x_{2} + u_2 - w_2)\rbrace = 0.1 J_3(x_{2} + u_2 - 0) + 0.2 J_3(x_{2} + u_2 - 1) + 0.2 J_3(x_{2} + u_2 - 2) + 0.3 J_3(x_{2} + u_2 - 3)
\end{equation}
which will result equal $-6-4.35=-10.35$. We get the values for $J_3$ from Table \ref{tab:k3}. The values for $J_2(x_2)$ can be seen in Table \ref{tab:k2}
\begin{table}[h]
	\centering
	\caption{k=2, the minimi per row is marked with bold}
	\label{tab:k2}
	\begin{tabular}{l|llll|l}
		k=2     & $u_2=0$         & $u_2=1$ & $u_2=2$        & $u_2=3$        & $\mu_2^*$ \\ \hline
		$x_2=0$ & 0               & -1.85   & -3.4           & \textbf{-4.35} & 3         \\
		$x_2=1$ & -5.85           & -6.4    & \textbf{-7.35} & -              & 2         \\
		$x_2=2$ & \textbf{-10.4}  & -10.35  & -              & -              & 0         \\
		$x_2=3$ & \textbf{-14.35} & -       & -              & -              & 0        
	\end{tabular}
\end{table}
For $k=1$ we only have to count the row for when $x_1=0$ sine we have no packages in the beginning. We will however count all, since we will need it in the b) part. The result can be seen from Table \ref{tab:k1}

\begin{table}[h]
	\centering
	\caption{k=1, the minimi per row is marked with bold}
	\label{tab:k1}
	\begin{tabular}{l|llll|l}
		k=1     & $u_1=0$         & $u_1=1$ & $u_1=2$        & $u_1=3$        & $\mu_1^*$ \\ \hline
		$x_1=0$ & -4.35           & -4.95   & -5.76          & \textbf{-6.17} & 3         \\
		$x_1=1$ & -8.95           & -8.76   & \textbf{-9.17} &        -        & 2         \\
		$x_1=2$ & \textbf{-12.76} & -12.17  &   -             &        -        & 0         \\
		$x_1=3$ & \textbf{-16.17} &   -      &      -          &       -         & 0        
	\end{tabular}
\end{table}
We can see that the optimal control policy is $\mu_1^*(x_1) = 3$ since $x_1=0$. Which gives us $J_1(0) = -6.17$\euro loss i.e. $6.17$\euro profit. The optimal control policies and expected profits for specific days are seen in each table.


\subsection*{b)}
We now have an additional day which leads to another iteration. Since the probability distribution is the same as for $k=1$ we can do the exactly same calculation except using $J_1$ isntead of $J_2$. Since we only have one possibility for $x_0$ which is $0$ we get the result seen in Table \ref{tab:k0}. The control policy is $\mu_0^*(0) = 3)$, which gives us $J_0(0) = -8.21$\euro loss i.e. $8.21$\euro profit.

\begin{table}[h]
	\centering
	\caption{k=0, the minimi per row is marked with bold}
	\label{tab:k0}
	\begin{tabular}{l|llll|l}
		k=0     & $u_0=0$         & $u_0=1$ & $u_0=2$        & $u_0=3$         & $\mu_0^*$ \\ \hline
		$x_0=0$ & -6.17           & -6.77   & -7.688         & \textbf{-8.206} & 3     
	\end{tabular}
\end{table}
\end{document}




